---
title: "Proposal title"
subtitle: "Proposal"
author: 
  - name: "Team Name"
    affiliations:
      - name: "School of Information, University of Arizona"
description: "Project description"
format:
  html:
    code-tools: true
    code-overflow: wrap
    code-line-numbers: true
    embed-resources: true
editor: visual
code-annotations: hover
execute:
  warning: false
---

```{r}
#| label: load-pkgs
#| message: false

if (!require("pacman")) 
  install.packages("pacman")

pacman::p_load(tidyverse,
              ggplot2)
```

## Datasets

```{r}
#| label: load-dataset
#| message: false

olympics <- read.csv("data/athlete_events.csv")
```

## **Introduction and data**

### Olympic Data Set (athlete_events.csv)

Our data set comes from 2024-08-06 tidytuesday, and contains data from every Olympic Games ranging from Athens 1896 to Rio 2016. The data set is 271,116 rows by 15 columns and contains data for every Olympic medal winner such as their name, sex, country, height, weight, and games they won the medal at.

It is important to note that winter and summer games were held in the same year up until 1992. From 1992 onward they were staggered by two years.

### \*Data set 2

\*Data set 2 description

### \* Data set 3

\* data set 3 description

## Questions

### Olympic Data Set (athlete_events.csv

1.  Is there an ideal height or weight for each event? If so, what are they and have they changed over time?
2.  

## Analysis plan

-   A plan for answering each of the questions including the variables involved, variables to be created (if any), external data to be merged in (if any).

## **Glimpse of data**

```{r}
glimpse(olympics)
```
