---
title: "Proposal title"
subtitle: "Proposal"
author: 
  - name: "Mint Minds For DataViz"
    affiliations:
      - name: "School of Information, University of Arizona"
description: "Project description"
format:
  html:
    code-tools: true
    code-overflow: wrap
    code-line-numbers: true
    embed-resources: true
editor: visual
code-annotations: hover
execute:
  warning: false
---

```{r}
#| label: load-pkgs
#| message: false

if (!require("pacman")) 
  install.packages("pacman")

pacman::p_load(tidyverse,
              ggplot2)
```

## Datasets

```{r}
#| label: load-dataset
#| message: false

olympics = read.csv("data/athlete_events.csv")
world_events = read.csv("data/world_important_dates.csv")
stock_prices = read.csv("data/big_tech_stock_prices.csv")
```

## **Introduction and data**

### Olympic Data Set (athlete_events.csv)

Our data set comes from 2024-08-06 tidytuesday, and contains data from every Olympic Games ranging from Athens 1896 to Rio 2016. The data set is 271,116 rows by 15 columns and contains data for every Olympic medal winner such as their name, sex, country, height, weight, and games they won the medal at.

It is important to note that winter and summer games were held in the same year up until 1992. From 1992 onward they were staggered by two years.

### World Events Data Set (world_important_events.csv)

This data set comes from Saket Kumar on Kaggle and will be used in conjunction with the Olympics Data Set. The data set is 1,096 rows by 12 columns and contains data about major historic events such as date, country, type of event and more.

### Big Tech Stock Prices (big_tech_stock_prices.csv)

Bit tech stock price dataset from 2023-02-07 tidytuesday. The data has 45089 rows and 8 columns. Rows has the stock stock symbol, date, open, high, low, close, adj close, volume from 2010-01-04 to 2022-12-29 of the following companies: Adobe Inc., Amazon.com, Inc., Salesforce, Inc., Cisco Systems, Inc., Alphabet Inc., IBM, Intel Corporation, Meta Platforms, Inc., Microsoft Corporation, Netflix, Inc., NVIDIA Corporation, Oracle Corporation, Tesla, Inc.

\*Data set 2 description

### \* Data set 3

\* data set 3 description

## Questions

### Olympic Data Set (athlete_events.csv)

1.  Is there an ideal height or weight for each event? If so, what are they and have they changed over time?
2.  How has the performance of different countries changed over time? Are there any major events that could have led to a change in performance?

### Big Tech Stock Prices (big_tech_stock_prices.csv)

1.  Is there any relationship between daily volatility open price and close price with the volume of stocks?
  - This question is important because it can help trader to understand the relationship between the volume of stocks and the daily volatility of the stock price.
2.  do closing price of stock of Big Tech companies correlate with each other over time?
  - This question is important because it provides traders insights of closing prices of Big Tech companies related to volume of stock, it will help them to make investment decisions.


### Data Set 3

1.  Q1
2.  Q2

## Analysis plan

-   A plan for answering each of the questions including the variables involved, variables to be created (if any), external data to be merged in (if any).

### Olympic Data Set (athlete_events.csv)

1.  Variables: Sex, Height, Weight, Medal, Sport, Games, Year. We will investigate the distributions of height and weight of athletes across different sports and sexes. We can also investigate if so-called "ideal" heights and weights have changed over time for each event. This question has the potential to raise ethical concerns in that anyone is able to succeed in any sport no matter their height/weight. We just want to investigate whether or not there is a historical trend of success across different events in athletes with certain heights/weights.
2.  Olympics data variables: Year, Games, Country, Medal, Event. World Events data variables: Name, Year, Month, Country, Type, Place, Impact, Effected Population, Person/Group Responsible, Outcome. We will investigate the change in medal count per country over time; overall and by event. We can use the world events data in order to find a potential cause for observed medal count changes.

### Big Tech Stock Prices (big_tech_stock_prices.csv)

1 . Variables: Date, Open, High, Low, Close, Adj Close, Volume. We will create new variables for daily volatility and investigate the relationship between daily volatility, open price, close price, and volume of stocks.
2.  Variables: Date, Close. We’ll examine how the closing price of different Big Tech companies moves together by calculating correlations between them.

### Data Set 3

## **Glimpse of data**

```{r}
glimpse(olympics)
glimpse(world_events)
glimpse(stock_prices)
```
