---
title: "Proposal title"
subtitle: "Proposal"
author: 
  - name: "Team Name"
    affiliations:
      - name: "School of Information, University of Arizona"
description: "Project description"
format:
  html:
    code-tools: true
    code-overflow: wrap
    code-line-numbers: true
    embed-resources: true
editor: visual
code-annotations: hover
execute:
  warning: false
---

```{r}
#| label: load-pkgs
#| message: false
library(tidyverse)
```

## Dataset

```{r}
#label: load-dataset
# message: false

# Option 1: tidytuesdayR package 
install.packages("tidytuesdayR")

tuesdata <- tidytuesdayR::tt_load('2024-08-06')
## OR
#tuesdata <- tidytuesdayR::tt_load(2024, week = 32)

olympics <- tuesdata$olympics


# Option 2: Read directly from GitHub

#olympics <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2024/2024-08-06/olympics.csv')

```

## **Introduction and data**

For each data set:

-   **Identify the source of the data.**

The olympic data

-   **State when and how it was originally collected (by the original data curator, not necessarily how you found the data).**

-   **Write a brief description of the observations.**

-   **Address ethical concerns about the data, if any.**

Our data set comes from 2024-08-06 tidytuesday, and contains data from every Olympic Games ranging from Athens 1896 to Rio 2016. The data set is 271,116 rows by 15 columns and contains data for every Olympic medal winner such as their name, sex, country, height, weight, and games they won the medal at.

It is important to note that winter and summer games were held in the same year up until 1992. From 1992 onward they were staggered by two years.

## Questions

The two questions you want to answer.

## Analysis plan

-   A plan for answering each of the questions including the variables involved, variables to be created (if any), external data to be merged in (if any).
