---
title: "Proposal title"
subtitle: "Proposal"
author: 
  - name: "Team Name"
    affiliations:
      - name: "School of Information, University of Arizona"
description: "Project description"
format:
  html:
    code-tools: true
    code-overflow: wrap
    code-line-numbers: true
    embed-resources: true
editor: visual
code-annotations: hover
execute:
  warning: false
---

```{r}
#| label: load-pkgs
#| message: false

if (!require("pacman")) 
  install.packages("pacman")

pacman::p_load(tidyverse,
              ggplot2)
```

## Datasets

```{r}
#| label: load-dataset
#| message: false

olympics <- read.csv("data/athlete_events.csv")
```

## **Introduction and data**

### Olympic Data Set (athlete_events.csv)

Our data set comes from 2024-08-06 tidytuesday, and contains data from every Olympic Games ranging from Athens 1896 to Rio 2016. The data set is 271,116 rows by 15 columns and contains data for every Olympic medal winner such as their name, sex, country, height, weight, and games they won the medal at.

It is important to note that winter and summer games were held in the same year up until 1992. From 1992 onward they were staggered by two years.

### Big Tech Stock Prices (big_tech_stock_prices.csv)

\*Data set 2 description

### \* Data set 3

\* data set 3 description

## Questions

### Olympic Data Set (athlete_events.csv)

1.  Is there an ideal height or weight for each event? If so, what are they and have they changed over time?
2.  How has the performance of different countries changed over time? Are there any major events that could have led to a change in performance?

### Big Tech Stock Prices (big_tech_stock_prices.csv)

1.  Q1
2.  Q2

### Data Set 3

1.  Q1
2.  Q2

## Analysis plan

-   A plan for answering each of the questions including the variables involved, variables to be created (if any), external data to be merged in (if any).

### Olympic Data Set (athlete_events.csv)

1.  Variables: Sex, Height, Weight, Medal, Sport, Games, Year. We will investigate the distributions of height and weight of athletes across different sports and sexes. We can also investigate if so-called "ideal" heights and weights have changed over time for each event. This question has the potential to raise ethical concerns in that anyone is able to succeed in any sport no matter their height/weight. We just want to investigate whether or not there is a historical trend of success across different events in athletes with certain heights/weights.

### Big Tech Stock Prices (big_tech_stock_prices.csv)

### Data Set 3

## **Glimpse of data**

```{r}
glimpse(olympics)
```
