# Visualization 1

```{r}
#| label: Visualization 1
#| warning: False
#| message: False

# Visualization 1

# Load in packages
if (!require("pacman")) 
  install.packages("pacman")

pacman::p_load(tidyverse,
               ggridges,
               here,
               RColorBrewer,
               DescTools,
               grid,
               caTools,
               pROC)


#### Load in Data
olympics <- read_csv(here("data", "athlete_events.csv"))

#### Data Prep

# create decade column, years are rounded down for decade column (EX: Year=1896,Decade=1890)
olympics$Decade <- floor(olympics$Year / 10) * 10

# create height in inches column
olympics$Height_Inches <- round(olympics$Height / 2.54)

# create new df containing only basketball data
basketball <- subset(olympics, Sport == 'Basketball')

# create height in feet column
basketball$Height_Ft <- floor(basketball$Height_Inches / 12)

# create remainder of inches column
basketball$Height_Inches_Rem <- round(basketball$Height_Inches %% 12)

# filter even further for only mens basketball data
mens_basketball <- subset(basketball, Event %in% c("Basketball Men's Basketball"))

# convert decade column to factor for plotting
mens_basketball$Decade = as.factor(mens_basketball$Decade)

# find mode for 1930 and 2010
basketball_summary_stats_by_decade = mens_basketball |>
  group_by(Decade) |>
  summarize(mode = Mode(Height_Inches, na.rm = TRUE)) |>
  ungroup()

mens_basketball|>
  ggplot(aes(x = Height_Inches, y = Decade, fill = Decade)) +
  geom_density_ridges() +
  scale_fill_brewer(palette = "Oranges") +
  scale_x_continuous(breaks = seq(60, 90, by = 5)) +
  labs(x = "Height (inches)",
       y = NULL,
       title = "The Increase of Olympic Men's Basketball Player Heights",
       subtitle = "from the 1930's to the 2010's") +
  geom_segment(aes(x = 80.65, y = 10.45, xend = 88, yend = 4.5), linewidth = 0.3) +
  geom_segment(aes(x = 72.2, y = 2.7, xend = 88, yend = 4.5), linewidth = 0.3) +
  geom_text(aes(x = 81.3, y = 10.8, label = "6'8\""), size = 3) +
  geom_text(aes(x = 72.7, y = 3.05, label = "6'0\""), size = 3) +
  annotate("point", x = 80.65, y = 10.45, fill = "red", size = 3, shape = 21, stroke = 0.5) +
  annotate("point", x = 72.2, y = 2.7, fill = "red", size = 3, shape = 21, stroke = 0.5) +
  geom_label(aes(x = 88, y = 4.5, label = "The most common\nplayer height has\nincreased by 8 inches\nsince the 1930's"),
             color = "black", fill = "white", size = 3, fontface = "bold") +
  theme(legend.position = "none",
        panel.background = element_rect(fill = "white"),
        plot.title = element_text(face = "bold"),
        axis.ticks.x = element_blank(),
        axis.ticks.y = element_blank()) +
  coord_cartesian(clip = "off")
```

## Viz 1 References

<https://r-coder.com/subset-r/>

<https://stackoverflow.com/questions/35352914/floor-a-year-to-the-decade-in-r>

# Visualization 2

Logistic Regression Model

```{r}
#| label: Logistic Regression Viz 2
#| warning: False

# Add binary medal column: 1 = medal, 2 = no medal
mens_basketball = mens_basketball |>
  mutate(medal_bi = case_when(
    Medal == "Bronze" ~ 1,
    Medal == "Silver" ~ 1,
    Medal == "Gold" ~ 1,
    is.na(Medal) ~ 0
  ))

# Create 2 column df of height and medal_bi and remove NA's
logit_df = as.data.frame(cbind(mens_basketball$Height_Inches,
                               mens_basketball$medal_bi))
colnames(logit_df) = c("height",
                       "medal")
logit_df = logit_df |>
  drop_na(height)

# Split into training and testing
set.seed(123)
split = sample.split(logit_df$medal, SplitRatio = 0.7)
train_data = subset(logit_df, split == TRUE)
test_data = subset(logit_df, split == FALSE)

# Fit the model
medal_logit = glm(medal ~ height, data = train_data, family = binomial)
summary(medal_logit)

# Test the model
test_data$pred_prob = predict(medal_logit, newdata = test_data, type = "response")

roc_curve = roc(test_data$medal, test_data$pred_prob)
plot(roc_curve)
auc(roc_curve)

# Plot the probabilities
height_seq = as.data.frame(seq(min(logit_df$height) - 20, max(logit_df$height) + 20, length.out = 100))
colnames(height_seq) = "height"
pred_prob2 = predict(medal_logit, newdata = height_seq, type = "response")

plot_data = as.data.frame(cbind(height_seq,
                                pred_prob2))

plot_data |>
  ggplot(aes(x = height, y = pred_prob2)) +
  geom_line(size = 1) +
  geom_jitter(data = test_data, aes(x = height, y = pred_prob, color = as.factor(medal)),
              width = 0, height = 0.1, alpha = 0.6) + 
  scale_color_manual(
    values = c("1" = "darkorange", "0" = "steelblue"),
    labels = c("Non-Medal Winner", "Medal Winner"),
    name = "Medal Status"
  ) +
  labs(
    title = "Probability of Winning an Olympic Medal by Height",
    subtitle = "Logistic Regression Model",
    x = "Height (in)",
    y = "Predicted Probability"
  ) + 
  scale_y_continuous(
    labels = scales::percent_format(),
    breaks = seq(0, 1, 0.20)
  ) +
  theme_minimal() + 
  theme(
    panel.grid = element_blank(),
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5, size = 12),
    axis.title = element_text(size = 12),
    legend.position = "right",
    legend.text = element_text(size = 10),
    legend.title = element_text(size = 11, face = "bold"),
    legend.background = element_rect(color = "black"),
    aspect.ratio = 1/2
  )
```

p-value = 6.05e-10 -\> very statistically significant

AUC = 0.5973 -\> just slightly better than random guessing

\^Together these values mean that height and medal are strongly correlated, but the model is not good at predicting based on height alone.

Height Coef = 0.08477

Odds Ratio = e\^0.08477 = 1.0885

For every one inch increase in height, the odds of winning a medal increase by 8.85%

Conclusion: Height almost definitely plays a role in earning a medal in Olympic basketball, but there are many other important factors needed to accurately predict if a player will win a medal such as skills, team context, or other physical attributes.

## Viz 2 References

<https://www.rpubs.com/fangya/logr>

<https://quarto.org/docs/dashboards/> (dashboard ref)

<https://plotly.com/r/>
